### 2015-03-07 sandbox/normal_approximation.tex
moved to sandbox

\documentclass{article}

\begin{document}

Model 
\begin{equation} 
Y_{gvi} \stackrel{ind}{\sim} NB\left(e^{\mu_{gv}+\gamma_{vi}},e^{\psi_g}\right), 
\label{e:data}
The probability mass function is 
\[ p(y_{gvi}|\mu_{gv},\gamma_{vi},\psi_g) = \frac{}{} \]

\end{equation}
with $E[Y_{gvi}] = \eta_{gvi} =e^{\mu_{gv}+\gamma_{vi}}$ and $V[Y_{gvi}] = \eta_{gvi} + e^{\psi_g}\eta_{gvi}^2$.
\[ \phi_g = \frac{\mu_{g1}+\mu_{g2}}{2}, \quad 
\alpha_g = \frac{\mu_{g1}-\mu_{g2}}{2}, \quad \mbox{and} \quad
\delta_g = \phi_g-\mu_{g3}. \]
Then the likelihood is 
\[ L(\phi_g,\alpha_g,\delta_g,\psi_g) \propto \]

\end{document}