### 2015-03-09 sandbox/integration.tex
methods of integration

\documentclass{article}

\usepackage{amsmath}

\newcommand{\argmax}{\mbox{argmax}}

\begin{document}

\section{MAP estimate}

Let $\theta_g$ be the gene-specific parameters and $\pi$ be the hyperparameters. The MAP estimate for $\pi$ is 
\begin{align*}
\argmax_{\pi} \log p(\pi|y) 
&= \argmax_{\pi} \log\left( \int \prod_{g=1}^G \left[ \, \prod_{v=1}^V \prod_{i=1}^{n_v} NB(y_{gvi}|f(\theta_g)) p(\theta_g|\pi) \right] p(\pi) d\theta \right) \\
&= \argmax_{\pi}\sum_{g=1}^G \log \left[ \, \int \prod_{v=1}^V \prod_{i=1}^{n_v} NB(y_{gvi}| f(\theta_g)) p(\theta_g|\pi) d\theta_g  \right] + \log(p(\pi)) \\
&= \argmax_{\pi}\sum_{g=1}^G \log(I_g) + \log(p(\pi)) 
\end{align*}
where 
 \[
 I_g = \int \prod_{v=1}^V \prod_{i=1}^{n_v} NB(y_{gvi}| f(\theta_g)) p(\theta_g|\pi) d\theta_g
 \]

\subsection{Quadrature}

Using quadrature, we would have a set of points $\theta_g^{(k)}$ that would depend on $\pi$ and associated weight $w_k$. We would approximate the integral via 
\[
I_g \approx \sum_{k=1}^K w_k NB(y_{gvi}| f(\theta_g^{(k)})).
\]

\subsection{Monte Carlo integration}

Alternatively, we can use Monte Carlo integration which would sample $\theta_g^{(k)} \sim p(\theta_g|\pi)$ and approximate the integral via 
\[
I_g \approx \frac{1}{K} \sum_{k=1}^K NB(y_{gvi}| f(\theta_g^{(k)})).
\]

\subsection{Importance sampling}

Importance sampling uses an alternative sampling distribution $q(\theta_g|\pi)$. Then we sample $\theta_g^{(k)} \sim q(\theta_g|\pi)$ and approximate the integral via 
\[
I_g \approx \sum_{k=1}^K w_k NB(y_{gvi}| f(\theta_g^{(k)})).
\]
where $w_k = p(\theta^{(k)}|\pi)/p(\theta^{(k)}|\pi)$.

If we make the proposal gene specific, i.e. $q_g(\theta_g|\pi)$, then we can sample using a proposal that is tuned to the data. For example, one component of $\theta_g$ is $\phi_g$, the mean parental response, i.e. $E[y_{g1i}+y_{g2i}]/2$. This parameter is extremely variable across $g$. It certainly makes sense to choose $\phi_g$ in a range near the empirical parental average, i.e.

\[ \hat{\phi}_g = \frac{1}{2}\left[ \frac{1}{n_1} \sum_{i=1}^{n_1} y_{g1i} + \frac{1}{n_2} \sum_{i=1}^{n_2} y_{g2i} \right] \]

\end{document}