### 2015-06-17 est_testing/prior_het_prob.tex
moved file to proper location

\documentclass{article}
\usepackage{amsmath,amsfonts}
\usepackage{setspace}
\begin{document}

\doublespace

Calculating prior probability of heterosis when
\[\alpha \sim Laplace(0,\sigma_\alpha),\; \delta \sim Laplace(0,\sigma_\delta)\]

First, note that $X = |\alpha| \sim Exp(\sigma_\alpha)$ and $Y = |\delta| \sim Exp(\sigma_\delta)$. Note also the pivotal quantities $X/\sigma_{\alpha}, Y/\sigma_{\delta} \sim Exp(1)$. Since we interested in the probability of the event $\{Y > X\} = \{Y/X>1\}$, consider the pivotal quantity, $Z = \frac{Y/\sigma_\delta}{X/\sigma_\alpha}$. By identifying $\frac{\sigma_\alpha}{\sigma_\delta} = c$, it is clear that $P(Z>c) = P(X/Y > 1)$.We can calculate the density of $Z$ by the transformation method as follows:

\[f_Z(z) = \int t\cdot f_{X,Y}(t,zt) dt = \int  t\cdot e^{-t(z+1)} dt = \frac{1}{(z+1)^2},\]

where the last equality comes from recognizing the integrand as the kernel of a gamma distribution.\\

Integrating this density, we find that $P(Z>c) = 1/(c+1)$. Given a probability $p$, $c$ can be calculated by $c = 1/p - 1$.\\

\end{document}