### 2015-09-16 supp.tex
add supplementary material

\documentclass[10pt]{article}

\usepackage{fullpage, hyperref}

\usepackage{amsmath}
\usepackage[figuresright]{rotating}

\def\bSig\mathbf{\Sigma}
\newcommand{\VS}{V\&S}
\newcommand{\tr}{\mbox{tr}}

\newcommand{\Stan}{{\tt Stan}}
\newcommand{\RStan}{{\tt RStan}}
\newcommand{\edgeR}{{\tt edgeR}}
\newcommand{\baySeq}{{\tt baySeq}}
\newcommand{\ShrinkBayes}{{\tt ShrinkBayes}}
\newcommand{\RNAseq}{RNA-seq}

%\renewcommand{\gamma}{c}

\title{Supplmentary material: Empirical Bayes analysis of \RNAseq{} data for detection of gene expression heterosis}

\author{Jarad Niemi, 
Eric Mittman, 
Will Landau, and 
Dan Nettleton \\
Department of Statistics, Iowa State University, Ames, Iowa, U.S.A.}

\begin{document}

\maketitle

The data and code necessary to reproduce the analyses in the manuscript are available at \url{https://github.com/jarad/niemi2015empirical}. 
The repository contains a {\tt README.md} file that describes the directory structure and files within those directories.
In particular, the data used in the real data analysis are available at \url{https://github.com/jarad/niemi2015empirical/tree/master/real/data}. 
If these data or code are used in a future publication, please cite the manuscript associated with this supplementary material.

\end{document}